On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-11 LME-MPM.tex
Cambios Miguel

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging
  under this approach due \textcolor{red}{to} the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter,
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements through numerical examples.    
\end{abstract}

@@

\setlength{\parskip}{1cm plus 5mm minus 4mm}
%\hline
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\large
\textbf{Highlights}
\normalsize
\begin{itemize}
\item We have proposed a new and efficient time integration scheme for the \acrshort{mpm} in order to dissipate spurious stress oscillations without damping the solution.
\item The proposal of the \acrshort{lme} approximation technique
  under the \acrshort{mpm} framework is a robust substitute of the
  traditional shape functions.
\item Our approach have clear benefits on large strain fast dynamic simulations, where the grid-crossing pathology appears. 
\end{itemize}

@@

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretisation by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

The article is organised as follows : Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the explicit predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems}, applications to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention: \textcolor{blue}{Any vectorial field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are used within paper. The subscript $\square_p$ is used to define a particle
variable. While the subscript $\square_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\square^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\square}$ and
$\ddot{\square}$ for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ means the
first and second contraction of a tensor, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{blue}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and an appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also denominated particles the manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be approximated within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}

@@

As stated previously, an explicit forward Euler algorithm has been utilized widely within the \acrshort{mpm} methodology. This scheme has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and can be sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Other authors have proposed many others time integration alternatives
like \cite{Guilkey_2003,Charlton_2017,Tran2019e}. In the
first publication on \acrshort{mpm} \cite{Sulsky1994}, \textcolor{blue}{the nodal acceleration
was considered to update the particles by}
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{a}_{I}^{k},
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} pointed out, this algorithm has been shown to be numerically unstable due to \textcolor{red}{the fact} that nodal forces can be infinite for an infinitesimal nodal mass $\tens{m}_I$. This issue may lead to numerical problems when nodal acceleration is obtained in the evaluation of the Eqs. \eqref{eq:Sulsky-1994-UL-x} and \eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm was proposed by Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}:
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{p}_{I}^{k}}{\tens{m}_I}, 
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{f}_{I}^{k}}{\tens{m}_I}.
\end{equation}
Delving into the improvement of the accuracy of the \acrshort{mpm} explicit schemes, Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for \acrshort{mpm} inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both \textcolor{red}{at} the beginning and \textcolor{red}{at} the end of the time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\  N_{Ip}\ \left[(1 - \gamma)\ \vec{a}_I^{k} +
    \gamma\ \vec{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} N_{Ip} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right],
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vec{a}_p^{k+1} = \sum_{I\ \in\ \mathcal{B}_p} N_{Ip}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

\textcolor{blue}{In this section, an explicit predictor-corrector time integration
scheme is proposed, which is based on the Newmark central differences explicit scheme.} This method is devoted to solve  system of equations of the type
{\color{red}
\begin{equation*}
 \sum_{J\ \in\ \mathcal{B}} \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}
}
The nodal \acrshort{mpm} stage allows to apply this method
within the \acrshort{mpm} framework in a similar manner that the one
proposed by Tran~\textit{et al.}~\cite{Tran2019e}. By using the predictor \textcolor{blue}{step}  it is possible to calculate nodal velocities and update particles position employing nodal values
of velocity and acceleration.

@@

  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I}.

@@

\textcolor{blue}{In our proposed scheme, accelerations \eqref{eq:Update-lagrangian-pce-a} and velocities \eqref{eq:Update-lagrangian-pce-v} are computed with information coming from the predictor step. On the other hand, positions \eqref{eq:Update-lagrangian-pce-u} are evaluated with information coming from the corrector step in the velocity term, while acceleration term introduces information from the corrector. Therefore it shares clear similarities with the \textit{leapfrog scheme} \cite{Zhang_book_2016},} where position is not updated at full time step, but the velocity is updated at half time steps. Notice also that, with this approach, the calculation of nodal momentum values are not required. Due to its simplicity, the proposed scheme can be implemented with minor modifications over the standard forward Euler. The full implementation is summarized in the algorithm \ref{alg:the_npc_alg}.

@@

where $p_I$ \textcolor{blue}{stands for probability of the random variable outcomes. By interpreting these probabilities as the shape functions $N_I(\vec{x})$ of an approximation scheme, equation \eqref{eq:Shannon-entropy} can be regarded as a measure of the uncertainty of the approximation. Thus, according to Jaynes's principle of maximum entropy \cite{Jaynes1957},} the least-biased approximation scheme is
given by

>>>>

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under \sout{the} large
  \textcolor{green}{deformations} \sout{deformation regime}. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging
  under this approach due \textcolor{red}{to} the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \sout{ \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter,
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements through numerical examples.} \textcolor{green}{The performance of both improvements is validated by the high quality results of the numerical examples}    
\end{abstract}

@@

\setlength{\parskip}{1cm plus 5mm minus 4mm}
%\hline
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\large
\textbf{Highlights}
\normalsize
\begin{itemize}
\item \sout{We have proposed a} A new \textcolor{green}{accurate} and efficient time integration scheme for the \acrshort{mpm} is proposed. \sout{in order to dissipate spurious stress oscillations without damping the solution.}
\item The \sout{proposal of the} \acrfull{lme} approximation technique  \sout{under the} is a robust substitute of the traditional shape functions \textcolor{green}{within the} \acrshort{mpm} framework .
\item \sout{Our} \textcolor{green}{The proposed approach might improve} \sout{have clear benefits on}  large strain fast dynamic \acrshort{mpm} simulations \sout{, where the grid-crossing pathology appears.} as grid-crossing pathology are clearly mitigated.  
\end{itemize}

@@

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretisation by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an \sout{explicit} predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

The article is organised as follows : Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the \sout{explicit} predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems}, applications to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention: \textcolor{blue}{Any \sout{vectorial} \textcolor{green}{vector} field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are \sout{used within paper} \textcolor{green}{considered:} \sout{The} subscript $\square_p$ is used to define a particle
variable, \sout{. While the} subscript $\square_I$ is reserved \sout{in this notation} for \sout{denoting} nodal variables, \sout{And finally, the} superscript $\square^{\psi}$ involves a
virtual magnitude. \sout{For the operators, the convention is} \textcolor{green}{The convention adopted for the operators is as folows:} $\dot{\square}$ and
$\ddot{\square}$ \textcolor{green}{are considered} for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ \sout{means the first and second contraction of a tensor} \textcolor{green}{are the single and double contraction of tensor index}, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{blue}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ \sout{and} \textcolor{green}{where} $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
\sout{formulation} of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and \sout{an} appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also \sout{denominated} \textcolor{green}{known as} particles \textcolor{green}{in this} \sout{the} manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be approximated within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}

@@

As stated previously, an explicit forward Euler algorithm has been utilized widely within the \acrshort{mpm} methodology. This scheme has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} \textcolor{purple}{and can be sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Other authors have proposed many others time integration alternatives
like \cite{Guilkey_2003,Charlton_2017,Tran2019e}. In the
first publication on \acrshort{mpm} \cite{Sulsky1994}, \textcolor{blue}{the nodal acceleration
was considered to update the particles by}
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{a}_{I}^{k},
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ N_{Ip}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} pointed out, this algorithm has been shown to be numerically unstable due to \textcolor{red}{the fact} that nodal forces can be infinite for an infinitesimal nodal mass $\tens{m}_I$. This issue may lead to numerical problems when nodal acceleration is obtained in the evaluation of the Eqs. \eqref{eq:Sulsky-1994-UL-x} and \eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm was proposed by Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}:
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{p}_{I}^{k}}{\tens{m}_I}, 
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\ \frac{N_{Ip}\ \vec{f}_{I}^{k}}{\tens{m}_I}.
\end{equation}
Delving into the improvement of the accuracy of the \acrshort{mpm} explicit schemes, Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for \acrshort{mpm} inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both \textcolor{red}{at} the beginning and \textcolor{red}{at} the end of the time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vec{v}_p^{k+1} = \vec{v}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} \Delta t\  N_{Ip}\ \left[(1 - \gamma)\ \vec{a}_I^{k} +
    \gamma\ \vec{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vec{x}_p^{k+1} = \vec{x}_p^{k} + \sum_{I\ \in\ \mathcal{B}_p} N_{Ip} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right],
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vec{a}_p^{k+1} = \sum_{I\ \in\ \mathcal{B}_p} N_{Ip}\ \vec{a}_{I}^{k+1}.
\end{equation}

@@

\textcolor{blue}{In this section, an explicit predictor-corrector time integration
scheme is proposed, which is based on the Newmark central differences explicit scheme.} This method is devoted to solve  system of equations of the type
{\color{red}
\begin{equation*}
 \sum_{J\ \in\ \mathcal{B}} \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}
}
\sout{The nodal \acrshort{mpm} stage} \textcolor{green}{The presence of a background grid of nodes} allows to apply this \textcolor{green}{integration} method
within the \acrshort{mpm} framework in a similar manner that the one
proposed by Tran~\textit{et al.}~\cite{Tran2019e}. By using the predictor \textcolor{blue}{step}  it is possible to calculate nodal velocities and update particles position employing nodal values
of velocity and acceleration.

@@

  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k}}{\tens{m}_I}.

@@

\textcolor{blue}{In \sout{our} \textcolor{green}{the} proposed scheme, accelerations \eqref{eq:Update-lagrangian-pce-a} and velocities \eqref{eq:Update-lagrangian-pce-v} are computed with information coming from the predictor step. On the other hand, positions \eqref{eq:Update-lagrangian-pce-u} are evaluated with information coming from the corrector step in the velocity term, while acceleration term introduces information from the corrector. Therefore, \sout{it shares clear} \textcolor{green}{it can be appreciated some} similarities with the \textit{leapfrog scheme} \cite{Zhang_book_2016},} where position is not updated at full time step, but the velocity is updated at half time steps. Notice also that, with this approach, the calculation of nodal momentum values are not required. Due to its simplicity, the proposed scheme can be implemented with minor modifications over the standard forward Euler. The full implementation is summarized in the algorithm \ref{alg:the_npc_alg}.

@@

where $p_I$ \textcolor{blue}{stands for probability of the random variable outcomes. By interpreting these probabilities as the shape functions $N_I(\vec{x})$ of an approximation scheme, equation \eqref{eq:Shannon-entropy} can be regarded as a measure of the uncertainty of the approximation. Thus, according to Jaynes's principle of maximum entropy \cite{Jaynes1957},} the least-biased approximation scheme \sout{is} \textcolor{green}{can be} given by
